\documentclass[11pt,twocolumn,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{microtype}
\geometry{top=2.2cm, bottom=2.5cm, left=2.0cm, right=2.0cm, columnsep=0.7cm}

%% Theorem environments
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%% Notation macros
\newcommand{\N}{\mathcal{N}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\H}{\mathcal{H}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tick}{\tau_0}
\newcommand{\sep}{\mathrm{sep}}
\newcommand{\yield}{\mathrm{yield}}
\newcommand{\floor}{\beta}
\newcommand{\dcat}{d_{\mathrm{cat}}}
\newcommand{\norm}[1]{\left\|#1\right\|}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\cell}{\mathrm{cell}}
\newcommand{\Scal}{S}
\newcommand{\Lyap}{V}
\newcommand{\kap}{\kappa}
\newcommand{\pres}{P}

\title{\textbf{Network Yield and Computing Allocation:\\
A Unified Framework for Distributed Task Scheduling\\
via Thermodynamic Resolution Floors, Cell-Partition Control,\\
and Market Clearing}}

\author{Kundai Farai Sachikonye}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We develop a self-contained theory of distributed computing allocation in which every schedulable resource, every state observation, and every economic signal shares a single irreducible quantum $\tick > 0$: the \emph{compute-tick}. The framework rests on three foundations assembled into a unified object. First, a \emph{finite observer axiom} forces every monitor to return a cell rather than a point, establishing a resolution floor $\floor \geq \tick > 0$ that bounds observation precision absolutely. Second, a \emph{cell-partition scheduler} compiled from offline stability analysis governs task placement with a Lyapunov guarantee and a Banach contraction whose ratio is computable without runtime measurement. Third, a \emph{separation-cost functional} assigns to every execution slot a yield price equal to its marginal contribution to useful task progress, and the clearing map for those prices coincides with the yield-optimal assignment and with deterministic scheduler closure at the same fixed point.

The central result is a \emph{Three-way Equivalence}: yield-optimality, deterministic closure, and market clearing are one fixed point, coinciding because $\tick$ serves simultaneously as the physical resolution floor for observations, the algorithmic closure threshold for the scheduler, and the minimum lot size for the clearing market. From this equivalence we derive, as theorems rather than assumptions, that executor threads run at forced slot-optimal utilisation ratios, that tasks sort onto execution slots by comparative advantage in compute-ticks, and that a purely local, residual-work-descent routing law drains every task queue in finite time because the separation cost raised by a stalled queue is identical to the yield signal that summons an executor. A structural incorruptibility theorem guarantees that no monitoring payload can influence the scheduling action stream without traversing an explicit approval boundary. All proofs are given in full; no prior companion paper is required.
\end{abstract}

\tableofcontents

%% ===================================================================
\section{Introduction}
\label{sec:intro}
%% ===================================================================

Distributed computing systems schedule tasks onto processors, manage queues of pending work, observe system state, and route tasks between machines. Despite decades of theory---backpressure scheduling \cite{tassiulas1992}, fluid models \cite{neely2010}, fair queueing \cite{demers1990}, cluster managers \cite{hindman2011,verma2015}---no unified account explains simultaneously why a correct scheduler \emph{cannot} observe state with arbitrary precision, why the scheduler that maximises useful throughput is the same one that clears a natural market, and why both coincide with the fixed point of a pure local rule requiring no global coordination.

This paper provides that account. The central object is a \emph{computing network} $\G = (\N, \E)$: a finite set of \emph{queue nodes} $\N$ (buffers for pending tasks) and a finite set of \emph{execution slots} $\E$ (processors), each with a capacity and a cost. Tasks carry private completion targets; the objective is useful target-ward progress per unit compute-tick. The key insight is that a single positive quantum $\tick > 0$---the minimum schedulable unit of processor time---appears in three distinct roles:
\begin{itemize}[nosep]
\item \emph{Physical}: the resolution floor below which no monitor can distinguish system states;
\item \emph{Algorithmic}: the gap below which the scheduler cannot improve by reassigning a single task;
\item \emph{Economic}: the minimum lot size in the clearing market for execution capacity.
\end{itemize}

When these three roles are played by the same $\tick$, the three fixed-point conditions coincide and the Three-way Equivalence holds. When the roles are played by distinct constants, the equivalence degrades gracefully: the slack grows to the maximum of the three constants and the clearing price deviates from the separation cost by the algorithmic-economic split.

\paragraph{Organisation.} Section~\ref{sec:model} defines the computing network model and all primitive objects. Section~\ref{sec:floor} establishes the resolution floor theorem. Section~\ref{sec:monitor} develops the monitoring architecture: S-entropy state space, trajectory completion operator, and the monitor--control separation theorem. Section~\ref{sec:scheduler} presents the cell-partition scheduler with its stability and contraction theorems. Section~\ref{sec:separation} defines separation costs and the yield functional. Section~\ref{sec:equivalence} proves the Three-way Equivalence. Section~\ref{sec:liveness} proves liveness under backpressure coupling. Section~\ref{sec:incorruptibility} proves structural incorruptibility. Section~\ref{sec:discussion} situates the results in the literature. Section~\ref{sec:conclusion} concludes.

\paragraph{Prerequisites.} The paper is self-contained. Familiarity with basic real analysis (metric spaces, fixed-point theorems) and linear algebra (eigenvalues, positive-definite matrices) suffices.

%% ===================================================================
\section{The Computing Network Model}
\label{sec:model}
%% ===================================================================

\subsection{Primitive objects}

\begin{definition}[Computing network]
\label{def:network}
A \emph{computing network} is a tuple $\G = (\N, \E, c, \ell, \bar{v})$ where:
\begin{itemize}[nosep]
\item $\N = \{n_1, \ldots, n_K\}$ is a finite set of \emph{queue nodes} (buffers);
\item $\E = \{e_1, \ldots, e_M\}$ is a finite set of \emph{execution slots} (processors);
\item $c : \E \to \mathbb{Z}_{>0}$ assigns each slot a \emph{thread capacity} $c(e) \geq 1$;
\item $\ell : \E \to \R_{>0}$ assigns each slot an \emph{expected task duration} $\ell(e) > 0$;
\item $\bar{v} : \E \to \R_{>0}$ assigns each slot a \emph{maximum throughput rate} $\bar{v}(e) > 0$.
\end{itemize}
\end{definition}

\begin{definition}[Compute-tick]
\label{def:tick}
The \emph{compute-tick} $\tick > 0$ is the minimum indivisible unit of processor time. No scheduling decision, observation, or price can be expressed to a precision finer than $\tick$.
\end{definition}

\begin{remark}
In practice $\tick$ corresponds to the OS scheduler quantum, the minimum time-slice allocated by the kernel. Its existence is not an engineering choice but a consequence of the finite-observer axiom proved in Section~\ref{sec:floor}.
\end{remark}

\begin{definition}[Task]
\label{def:task}
A \emph{task} $x$ is a tuple $(o(x), \tau(x), w(x))$ where $o(x) \in \N$ is the \emph{originating queue node}, $\tau(x) \in \R^d$ is the \emph{completion target} (a feature vector encoding the desired output state), and $w(x) > 0$ is the \emph{work estimate} (expected compute-ticks to completion).
\end{definition}

\begin{definition}[Residual work]
\label{def:residual}
For a task $x$ currently executing on slot $e$, the \emph{residual work} is
\[
  r(x, e) \;=\; \rho\bigl(\mathrm{loc}(x),\; \tau(x)\bigr),
\]
where $\mathrm{loc}(x) \in \R^d$ is the current state of $x$'s computation (the partial result produced so far) and $\rho$ is the Euclidean distance on $\R^d$.
\end{definition}

\begin{definition}[Queue occupancy and directional pressure]
\label{def:pressure}
Let $\X(n)$ denote the set of tasks currently queued at node $n \in \N$, and let $\mathrm{hd}(e) \in \R^d$ denote the ``head'' of slot $e$: the state at which a task assigned to $e$ arrives after one tick of execution. The \emph{directional pressure} from node $n$ towards slot $e$ is
\[
  \pres(n, e) \;=\; \sum_{x \in \X(n)} \bigl[r(x, e) - \rho(\mathrm{hd}(e), \tau(x))\bigr]^{+},
\]
where $[z]^+ = \max(z, 0)$. Tasks that would \emph{not} make progress by moving to $e$ contribute zero pressure.
\end{definition}

\begin{definition}[Network transport yield]
\label{def:yield}
The \emph{network transport yield} of an assignment $A : \X \to \E$ (a map from tasks to slots) is
\[
  \yield(A) \;=\; \frac{\displaystyle\sum_{x \in \X} \bigl[r(x, A(x)) - \rho(\mathrm{hd}(A(x)), \tau(x))\bigr]^{+}}
                        {\displaystyle\sum_{e \in \E} \tick \cdot c(e) \cdot g_u(v(e))},
\]
where $v(e) \in (0, \bar{v}(e)]$ is the throughput rate at which slot $e$ operates under assignment $A$, and $g_u : \R_{>0} \to \R_{>0}$ is a strictly convex \emph{utilisation cost} satisfying $g_u' > 0$ and $g_u'' > 0$.

\noindent
The numerator counts total target-ward progress; the denominator counts total resource consumption in compute-ticks weighted by utilisation cost.
\end{definition}

\subsection{Separation cost}

\begin{definition}[Separation cost]
\label{def:sep}
The \emph{separation cost} of slot $e$ under assignment $A$ is the marginal yield lost by removing $e$ from the network:
\[
  \sep(e, A) \;=\; \yield(A) \;-\; \yield(A|_{\E \setminus \{e\}}),
\]
where $A|_{\E \setminus \{e\}}$ is the optimal re-assignment of tasks after removing $e$.
\end{definition}

\begin{remark}
$\sep(e, A) \geq 0$. A redundant slot ($e$ has a perfect substitute) has $\sep = 0$; a bottleneck slot has large $\sep$.
\end{remark}

\subsection{Assignments and closure}

\begin{definition}[Single-slot reassignment neighbourhood]
\label{def:neighborhood}
Two assignments $A$ and $A'$ are \emph{$\tick$-adjacent} if they differ in the allocation of at most one task and the resulting change in yield is at most $\tick$.
\end{definition}

\begin{definition}[Deterministic closure]
\label{def:closure}
Assignment $A$ is \emph{deterministically closed} if no $\tick$-adjacent assignment $A'$ satisfies $\yield(A') > \yield(A) + \tick$. Equivalently, the best single-task reassignment cannot improve yield beyond the resolution floor.
\end{definition}

%% ===================================================================
\section{The Resolution Floor}
\label{sec:floor}
%% ===================================================================

\subsection{Finite observer axiom}

\begin{assumption}[Finite observer]
\label{ass:finite}
Any monitoring subsystem of the computing network is itself a finite computational process: it runs on hardware with a finite number of distinguishable states, communicates over channels with finite bandwidth, and terminates its observation procedure in finite time.
\end{assumption}

This axiom is not a limitation of current technology; it is a logical consequence of a bounded physical substrate. It implies the following foundational theorem.

\begin{theorem}[Resolution floor]
\label{thm:floor}
Under Assumption~\ref{ass:finite}, there exists $\floor > 0$ such that for every observable quantity $q$ and every monitor $\mathcal{M}$, the identification $\mathcal{M}(q)$ satisfies
\[
  \inf_q \;\mathrm{diam}\bigl(\mathcal{M}(q)\bigr) \;\geq\; \floor \;>\; 0,
\]
where $\mathrm{diam}(\cdot)$ is the diameter of the set $\mathcal{M}(q) \subseteq \R$ returned by $\mathcal{M}$.

No monitor returns a point; every identification carries strictly positive irreducible residual uncertainty.
\end{theorem}

\begin{proof}
Let $S$ be the finite state space of $\mathcal{M}$ and let $Q$ be the range of $q$. $\mathcal{M}$ defines a function $f : Q \to 2^Q$. Since $S$ is finite, $f$ has at most $|S|$ distinct outputs. The pre-images $\{f^{-1}(r) : r \in \mathrm{range}(f)\}$ partition $Q$ into at most $|S|$ cells. Since $Q$ is a metric space with at least $|S|+1$ distinct points (for any finite $S$, the range of a physical quantity is infinite in principle), at least one cell has diameter $\geq \mathrm{diam}(Q)/|S| > 0$. Setting $\floor = \min_{r} \mathrm{diam}(f^{-1}(r))$ and noting that every cell must have positive diameter (because $q$ is a continuous physical quantity) gives $\floor > 0$.

Formally: suppose for contradiction that some cell $f^{-1}(r)$ has diameter zero. Then $f^{-1}(r)$ is a single point, and $f$ distinguishes that point from all others---which requires $|S| \geq |Q|$. But $Q$ is uncountable (it is a real-valued quantity) and $S$ is finite, a contradiction.
\end{proof}

\begin{corollary}[Compute-tick as floor]
\label{cor:tick_floor}
The compute-tick $\tick$ satisfies $\tick \geq \floor$. Scheduling decisions made at resolution $\tick$ are consistent with the observation resolution guaranteed by Theorem~\ref{thm:floor}.
\end{corollary}

\begin{proof}
Any scheduling decision that distinguishes two system states separated by less than $\floor$ would require a monitor that violates Theorem~\ref{thm:floor}. Therefore, the minimum indivisible decision granularity must be $\geq \floor$. Setting $\tick = \floor$ achieves the tightest consistent granularity.
\end{proof}

%% ===================================================================
\section{Monitoring Architecture}
\label{sec:monitor}
%% ===================================================================

\subsection{S-entropy state space}

\begin{definition}[S-normalisation]
\label{def:snorm}
For a physical quantity $q$ with known bounds $[q_{\min}, q_{\max}]$, the \emph{S-normalisation} is
\[
  \Scal(q) \;=\; \frac{q - q_{\min}}{q_{\max} - q_{\min}} \;\in\; [0,1].
\]
S-normalisation maps heterogeneous dimensional quantities onto a common unit interval, enabling uniform treatment of queue lengths, utilisation rates, task progress fractions, and thermal metrics.
\end{definition}

\begin{definition}[S-entropy state space]
\label{def:sspace}
Let the observable state of the computing network be described by $m$ independent quantities $q_1, \ldots, q_m$. The \emph{S-entropy state space} is
\[
  \H \;=\; [0,1]^m,
\]
equipped with the Euclidean metric. Each coordinate $\Scal(q_i)$ represents one observable dimension. $\H$ is compact, connected, and a complete metric space.
\end{definition}

\begin{remark}
The three canonical coordinates are: $s_k = \Scal(\text{queue occupancy})$ (knowledge entropy), $s_t = \Scal(\text{task age})$ (temporal entropy), and $s_e = \Scal(\text{residual work fraction})$ (evolution entropy). Additional coordinates may encode thermal load, network latency, and memory pressure.
\end{remark}

\subsection{Cell partition and the monitoring cell}

\begin{definition}[Cell partition]
\label{def:cellpartition}
A \emph{cell partition} $\C = \{C_1, \ldots, C_N\}$ is a finite Borel partition of $\H$ such that each cell $C_i$ has width $w_i \geq \floor > 0$.
\end{definition}

\begin{definition}[Monitoring cell]
\label{def:moncell}
For an observed state $h \in \H$, the \emph{monitoring cell} is the unique $C_i \in \C$ containing $h$:
\[
  \cell(h) \;=\; C_i \quad \Leftrightarrow \quad h \in C_i.
\]
The monitor returns $\cell(h)$, not $h$ itself. This is consistent with Theorem~\ref{thm:floor}: the cell has diameter $\geq \floor > 0$.
\end{definition}

\subsection{Trajectory completion}

Computing networks exhibit partially observable dynamics: a task's computational trajectory from initiation to completion is not directly accessible but must be inferred from snapshots of residual work.

\begin{definition}[Fuzzy network state]
\label{def:fuzzy}
A \emph{fuzzy network state} is a function $\mu : \N \times \E \to [0,1]$ assigning to each (queue-node, slot) pair a belief weight, with $\sum_{n,e} \mu(n,e) = 1$.
\end{definition}

\begin{definition}[State space of fuzzy states]
\label{def:fuzzyspace}
Let $\F$ be the set of all fuzzy network states. Equipped with the $\ell^1$ metric $d_1(\mu, \nu) = \sum_{n,e} |\mu(n,e) - \nu(n,e)|$, $(\F, d_1)$ is a complete metric space.
\end{definition}

\begin{definition}[Trajectory completion operator]
\label{def:trajop}
Define the \emph{trajectory completion operator} $\T = \T_{\mathrm{kin}} \circ \T_{\mathrm{therm}} \circ \T_{\mathrm{bal}}$ on $\F$:
\begin{itemize}[nosep]
\item $\T_{\mathrm{bal}}$ rebalances belief weights to be consistent with observed queue occupancies;
\item $\T_{\mathrm{therm}}$ applies a diffusion step consistent with the maximum-entropy principle over unobserved paths;
\item $\T_{\mathrm{kin}}$ projects forward by one tick using the assignment kinematics.
\end{itemize}
\end{definition}

\begin{theorem}[Trajectory completion converges]
\label{thm:traj}
Under Assumption~\ref{ass:finite} and with a fixed assignment $A$, the operator $\T$ is a Banach contraction on $(\F, d_1)$ with contraction ratio $\rho_{\T} < 1$. In particular, there exists a unique fixed point $\mu^* \in \F$ with $\T(\mu^*) = \mu^*$, and for any initial $\mu_0 \in \F$,
\[
  d_1\bigl(\T^k(\mu_0),\; \mu^*\bigr) \;\leq\; \rho_{\T}^k \cdot d_1(\mu_0, \mu^*),
\]
so the sequence $(\T^k(\mu_0))$ converges geometrically to $\mu^*$.
\end{theorem}

\begin{proof}
$\T_{\mathrm{bal}}$ is a projection onto the closed convex set of distributions consistent with observations; as a nearest-point projection in $\ell^1$ onto a convex set, it is non-expansive. $\T_{\mathrm{therm}}$ applies a strictly positive stochastic matrix $Q$ (the maximum-entropy diffusion kernel); by the Perron--Frobenius theorem, $Q$ has a unique stationary distribution and its spectral gap is $1 - \lambda_2(Q) > 0$, so multiplication by $Q$ is a contraction in $\ell^1$ with ratio $\leq 1 - \text{gap}(Q) < 1$. $\T_{\mathrm{kin}}$ is a deterministic permutation of the state; it is an isometry, hence non-expansive. The composition of a strict contraction with non-expansive maps is a strict contraction \cite{banach1922}. Setting $\rho_{\T} = 1 - \text{gap}(Q) < 1$ completes the proof.
\end{proof}

\subsection{Closure dichotomy for monitoring}

\begin{definition}[Monitoring confirmation]
\label{def:confirm}
A monitoring process over a finite set of sources $\Gamma = \{\gamma_1, \ldots, \gamma_n\}$ terminates in one of two states: \emph{confirmed} (sufficient evidence accumulated) or \emph{declined} (insufficient evidence; action withheld).
\end{definition}

\begin{theorem}[Confirm-or-decline dichotomy]
\label{thm:dichotomy}
Every monitoring process over finitely many sources with the closure criterion (Definition~\ref{def:closure} adapted to monitoring: no additional evidence reduces cell diameter below $\floor$) terminates in confirmed or declined. There is no third outcome.
\end{theorem}

\begin{proof}
The monitoring process accumulates evidence $e_1, e_2, \ldots$ from $n$ finite sources. By Assumption~\ref{ass:finite}, each source is a finite-state machine; the Cartesian product of $n$ finite-state machines is finite. The process therefore explores a finite directed graph of evidence states. Under the closure criterion, the process halts when no new transition reduces the diameter of the identified cell below $\floor$. Since the graph is finite, the process must reach a terminal state in finite steps. Terminal states partition into: states where the current cell meets the confirmation threshold (confirmed) and states where it does not (declined). There is no cycle that avoids both (the acyclicity follows from the requirement that each step must either reduce diameter or contribute new evidence; in a finite graph this eventually exhausts).
\end{proof}

\subsection{Coherence and sensor federation}

\begin{definition}[Mutually supporting cycle]
\label{def:cycle}
A set of monitoring sources $\{\gamma_1, \ldots, \gamma_n\}$ forms a \emph{mutually supporting cycle} if for each $i$, source $\gamma_i$ can confirm the output of $\gamma_{i-1 \bmod n}$ independently of all other sources.
\end{definition}

\begin{theorem}[Coherence requires three]
\label{thm:coherence}
A monitoring system is resilient to the arbitrary failure of any single source if and only if the remaining sources form a mutually supporting cycle of length $\geq 3$.
\end{theorem}

\begin{proof}
\emph{Necessity.} If all cycles in the dependency graph have length $\leq 2$, then every pair of sources supports only each other. Removing either element of such a pair destroys all support for the other, so the system is not resilient to single-source failure.

\emph{Sufficiency.} Suppose there exists a cycle $\gamma_1, \ldots, \gamma_n$ with $n \geq 3$. Remove source $\gamma_j$ for any $j$. The remaining cycle $\gamma_1, \ldots, \gamma_{j-1}, \gamma_{j+1}, \ldots, \gamma_n$ still has length $n-1 \geq 2$. For $n \geq 3$, $n - 1 \geq 2$ and the cycle is non-trivial. By the definition of mutual support, $\gamma_{j-1}$ is supported by $\gamma_{j-2}$ and $\gamma_{j+1}$ is supported by $\gamma_{j+2}$; the chain remains intact. Monitoring evidence can still propagate around the cycle, and the confirm-or-decline dichotomy (Theorem~\ref{thm:dichotomy}) is preserved.
\end{proof}

\begin{definition}[Multiplicative confirmation composition]
\label{def:multconf}
For independent confirmation sources $\gamma_1, \ldots, \gamma_n$ each with confirmation strength $\kap_i \in [0,1]$, the \emph{composite confirmation strength} is
\[
  \kap(\gamma_1 \circ \cdots \circ \gamma_n) \;=\; 1 - \prod_{i=1}^n (1 - \kap_i).
\]
\end{definition}

\subsection{Monitor--control separation theorem}

\begin{theorem}[Monitor--control separation]
\label{thm:separation}
A subsystem that acts on its own observations cannot simultaneously serve as an unbiased monitor of the system it controls.

Formally: let $\mathcal{M}$ be a monitor and $\mathcal{C}$ be a controller for the same system $\Sigma$. If $\mathcal{C}$'s actions are a function of $\mathcal{M}$'s outputs, then $\mathcal{M}$'s outputs are not independent of $\mathcal{C}$'s actions. In particular, any $\tick$-precise action by $\mathcal{C}$ induces a selection bias in $\mathcal{M}$ of at least $\floor$, violating the unbiasedness condition.
\end{theorem}

\begin{proof}
Let $h \in \H$ be the true system state. $\mathcal{M}$ returns $\cell(h)$ (Def.~\ref{def:moncell}). $\mathcal{C}$ reads $\cell(h)$ and executes action $u = A(\cell(h))$. The action $u$ modifies $\Sigma$ and therefore $h$: the post-action state is $h' = f(h, u)$ for some dynamics $f$. If $\mathcal{M}$ subsequently observes $h'$ to inform the same control cycle, its next output $\cell(h')$ depends on $u$, hence on $\cell(h)$, hence on the previous output of $\mathcal{M}$. The observation is therefore conditioned on the action, not independent of it.

For unbiasedness to hold, the monitor's output distribution must have conditional expectation $\mathbb{E}[\cell(h) | u] = \mathbb{E}[\cell(h)]$ for all $u$. But $u = A(\cell(h))$ makes $u$ a deterministic function of $\cell(h)$, so $h' = f(h, A(\cell(h)))$; the expectation $\mathbb{E}[\cell(h') | \cell(h)]$ is a non-trivial conditional distribution, not the marginal. The selection bias is the KL divergence between the conditional and marginal; since $\cell$ has diameter $\geq \floor > 0$ and $A$ maps distinct cells to distinct actions (by the schedulability condition), this divergence is at least $\log(1 + \floor / w_{\min}) > 0$, where $w_{\min}$ is the minimum cell width.

Therefore, monitoring and control must be separate processes whose outputs are never in the same feedback loop within one sampling period.
\end{proof}

%% ===================================================================
\section{Cell-Partition Scheduler}
\label{sec:scheduler}
%% ===================================================================

\subsection{Offline compilation}

The cell-partition scheduler pre-computes a lookup table $A : \C \to \E^{\X}$ mapping each monitoring cell to a task-to-slot assignment. This computation is performed offline; at runtime the scheduler reads the monitoring cell and performs a single table lookup.

\begin{definition}[Cell-partition scheduler]
\label{def:scheduler}
The \emph{cell-partition scheduler} is the function
\[
  u(t) \;=\; A\bigl(\cell(\Scal(h(t)))\bigr),
\]
where $h(t)$ is the observed state at time $t$, $\Scal$ applies S-normalisation (Def.~\ref{def:snorm}), $\cell$ returns the monitoring cell (Def.~\ref{def:moncell}), and $A$ is the pre-compiled assignment table.
\end{definition}

\begin{remark}
Because the cell partition has width $\geq \floor \geq \tick$, the scheduler operates at resolution consistent with Corollary~\ref{cor:tick_floor}. It never attempts to distinguish states separated by less than $\tick$.
\end{remark}

\subsection{Stability}

\begin{assumption}[Per-cell dynamics]
\label{ass:dynamics}
For each cell $C_i \in \C$, the closed-loop system under assignment $A(C_i)$ has linearised dynamics $\dot{h} = F_i(h)$ with a quadratic Lyapunov function $\Lyap_i(h) = h^T P_i h$ satisfying
\[
  F_i^T P_i + P_i F_i \;\preceq\; -\alpha_i P_i, \quad \alpha_i > 0.
\]
\end{assumption}

\begin{theorem}[Piecewise Lyapunov stability]
\label{thm:lyapunov}
Under Assumption~\ref{ass:dynamics} and with the cell-partition scheduler (Def.~\ref{def:scheduler}), the network state $h(t)$ converges globally asymptotically to the target cell $C^* \ni h^*$ containing the desired operating point.
\end{theorem}

\begin{proof}
Define the \emph{composite Lyapunov function}
\[
  \Lyap(h) \;=\; \Lyap_i(h) \quad \text{for } h \in C_i.
\]
On each cell $C_i$, $\dot{\Lyap}_i = h^T(F_i^T P_i + P_i F_i)h \leq -\alpha_i h^T P_i h = -\alpha_i \Lyap_i(h)$. At cell boundaries, the assignment switches from $A(C_i)$ to $A(C_j)$; by the cell-width condition ($w_i \geq \floor$), the boundary crossing requires a state change of magnitude $\geq \floor/2$, which is detectable. Across boundaries, $\Lyap$ may increase by at most $\Delta_{ij} = \sup_{h \in C_i \cap \partial C_j} |\Lyap_j(h) - \Lyap_i(h)|$. Under the Nyquist-type bound (Theorem~\ref{thm:nyquist}), the number of crossings per unit time is bounded, so the average $\Lyap$ decrease per crossing exceeds the boundary increment. By LaSalle's invariance principle applied piecewise, $h(t) \to C^*$ as $t \to \infty$.
\end{proof}

\begin{theorem}[Nyquist-type cell-width bound]
\label{thm:nyquist}
For the cell-partition scheduler to avoid aliasing (sampling faster than the cell permits), the cell width must satisfy
\[
  \floor \;\leq\; w_{\cell} \;\leq\; \frac{1}{2 B_{\mathrm{proc}}},
\]
where $B_{\mathrm{proc}}$ is the bandwidth of the fastest monitored process (inverse of the minimum observable period).
\end{theorem}

\begin{proof}
The lower bound $w_{\cell} \geq \floor$ is Corollary~\ref{cor:tick_floor}. The upper bound follows from the Nyquist--Shannon sampling theorem \cite{shannon1948}: to faithfully represent a signal of bandwidth $B_{\mathrm{proc}}$, the sampling interval must be $\leq 1/(2 B_{\mathrm{proc}})$. Since each cell corresponds to one sampling interval, $w_{\cell} \leq 1/(2 B_{\mathrm{proc}})$.
\end{proof}

\subsection{Discrete-time contraction}

In practice, the scheduler operates in discrete time with step $h = \tick$.

\begin{theorem}[Banach contraction for discrete scheduler]
\label{thm:contraction}
Let the discrete dynamics under $A(C_i)$ be $h_{t+1} = h_t + \tick F_i(h_t)$. Define the discrete operator $\T_i : C_i \to C_i$ by $\T_i(h) = h + \tick F_i(h)$. If $\tick < 2\alpha_i / L_i^2$ where $L_i$ is the Lipschitz constant of $F_i$ on $C_i$, then $\T_i$ is a Banach contraction with ratio
\[
  \rho_i \;=\; \sqrt{1 - 2\tick\alpha_i + \tick^2 L_i^2} \;<\; 1.
\]
\end{theorem}

\begin{proof}
For $h, h' \in C_i$:
\begin{align*}
  \norm{\T_i(h) - \T_i(h')}^2 &= \norm{(h - h') + \tick(F_i(h) - F_i(h'))}^2 \\
  &= \norm{h-h'}^2 + 2\tick\langle h-h', F_i(h) - F_i(h')\rangle \\
  &\quad + \tick^2\norm{F_i(h) - F_i(h')}^2.
\end{align*}
By Assumption~\ref{ass:dynamics}, $2\langle h-h', F_i(h) - F_i(h')\rangle \leq -2\alpha_i\norm{h-h'}^2$. By Lipschitz continuity of $F_i$, $\norm{F_i(h) - F_i(h')} \leq L_i\norm{h-h'}$. Therefore
\[
  \norm{\T_i(h) - \T_i(h')}^2 \leq (1 - 2\tick\alpha_i + \tick^2 L_i^2)\norm{h-h'}^2.
\]
Setting $\rho_i = \sqrt{1 - 2\tick\alpha_i + \tick^2 L_i^2}$, the condition $\rho_i < 1$ is equivalent to $\tick^2 L_i^2 - 2\tick\alpha_i < 0$, i.e., $\tick < 2\alpha_i/L_i^2$.
\end{proof}

\begin{corollary}[Unique fixed point per cell]
By the Banach fixed-point theorem \cite{banach1922}, each $\T_i$ has a unique fixed point $h_i^* \in C_i$, and the iteration $h_{t+1} = \T_i(h_t)$ converges geometrically.
\end{corollary}

\subsection{CUSUM--cell-crossing identity}

The cell-partition scheduler generates detectable events at cell boundaries. These boundary crossings are identical to out-of-control alarms in a CUSUM quality-control chart \cite{cusum1954}.

\begin{theorem}[CUSUM--cell-crossing identity]
\label{thm:cusum}
A cell boundary crossing by the scheduler (i.e., a transition from cell $C_i$ to cell $C_j$) is equivalent to a CUSUM out-of-control alarm with decision limit $H = w_{\cell}/2$. Conversely, every CUSUM alarm at limit $H = w_{\cell}/2$ corresponds to a cell boundary crossing.
\end{theorem}

\begin{proof}
Let $S_t = \sum_{k=1}^t (z_k - \mu_0)$ be the CUSUM statistic for observations $z_k$ about nominal mean $\mu_0$. A cell boundary at $\pm w_{\cell}/2$ from the cell centre corresponds to an alarm when $|S_t| > w_{\cell}/2$. The CUSUM statistic accumulates deviations from the cell centre; the alarm fires exactly when the accumulated deviation reaches the boundary. Conversely, the scheduler crosses the boundary exactly when the observed state $\Scal(h(t))$ exits $C_i$, which occurs when the S-normalised state deviates from $C_i$'s centre by more than $w_i/2 = w_{\cell}/2$. The two conditions are identical.
\end{proof}

%% ===================================================================
\section{Separation Costs and Network Yield}
\label{sec:separation}
%% ===================================================================

\subsection{Slot-level separation costs}

Recall Definition~\ref{def:sep}. We now establish properties of $\sep(e, A)$ that are essential for the Three-way Equivalence.

\begin{lemma}[Monotone reachability]
\label{lem:reach}
From any assignment $A$, the sequence of assignments generated by the greedy rule (always reassign the task that maximally increases yield, by at least $\tick$) is monotone non-decreasing in $\yield$ and terminates at a deterministically closed assignment in at most $|\X| \cdot |\E|$ steps.
\end{lemma}

\begin{proof}
Each step of the greedy rule increases $\yield$ by at least $\tick > 0$ (by the definition of the greedy rule; if no improvement of $\geq \tick$ exists, the rule does not fire and closure holds). The range of $\yield$ is bounded above by $\yield^* = \max_A \yield(A)$. Since each step increases yield by a discrete multiple of $\tick$, the total number of steps is at most $\lceil \yield^*/\tick \rceil < \infty$. More precisely, each task can be reassigned at most $|\E|$ times (once to each slot), so the total number of steps is $\leq |\X| \cdot |\E|$.
\end{proof}

\begin{proposition}[Separation cost as price]
\label{prop:sepcost}
At any assignment $A$, $\sep(e, A)$ is the correct clearing price for slot $e$ in the market of Definition~\ref{def:market}: a task's willingness to pay for slot $e$ is exactly the yield it gains by using $e$ rather than its next-best alternative.
\end{proposition}

\begin{proof}
Task $x$ gains yield $\Delta(x, e) = \yield(A) - \yield(A|_{x \not\to e})$ from using slot $e$. Summing over all tasks assigned to $e$, $\sum_x \Delta(x, e) = \sep(e, A)$ by definition. In the clearing market, the price clears supply and demand: the price $p(e)$ at which total demand (sum of willingness-to-pay) equals supply (capacity $c(e)$) is precisely $\sep(e, A)/c(e)$. For unit-capacity slots, $p(e) = \sep(e, A)$.
\end{proof}

\subsection{The yield market}

\begin{definition}[Yield market]
\label{def:market}
The \emph{yield market} is the assignment mechanism in which:
\begin{itemize}[nosep]
\item Each slot $e$ is offered at price $p(e)$;
\item Each task $x$ demands the slot $e^*(x) = \arg\max_{e} [\yield_x(e) - p(e)]$, where $\yield_x(e)$ is the yield contributed by assigning $x$ to $e$;
\item The market clears when supply equals demand for each slot.
\end{itemize}
\end{definition}

\begin{definition}[Market clearing]
\label{def:clearing}
A price vector $(p(e))_{e \in \E}$ and assignment $A$ form a \emph{market clearing} if:
\begin{enumerate}[nosep]
\item $A$ is individually rational: each task prefers its assigned slot to all others at the prevailing prices;
\item Markets clear: $\sum_{x: A(x)=e} 1 = c(e)$ for all $e$ (or the price is zero if supply exceeds demand);
\item Prices are non-negative: $p(e) \geq 0$.
\end{enumerate}
\end{definition}

%% ===================================================================
\section{The Three-Way Equivalence}
\label{sec:equivalence}
%% ===================================================================

This section proves the paper's central result.

\begin{assumption}[Positive utilisation cost]
\label{ass:cost}
The utilisation cost $g_u$ is strictly convex with $g_u(0) = 0$, $g_u' > 0$, $g_u'' > 0$, and $g_u(v) \to \infty$ as $v \to \bar{v}$.
\end{assumption}

\begin{assumption}[Gradient connectivity]
\label{ass:gradient}
For each task $x$ with target $\tau(x)$ and for each assignment $A$, there exists a sequence of single-slot reassignments from $A$ that strictly reduces $r(x, \cdot)$ at each step (i.e., the residual work descends monotonically along the reassignment sequence).
\end{assumption}

\begin{theorem}[Three-way Equivalence]
\label{thm:trinity}
Under Assumptions~\ref{ass:finite}, \ref{ass:cost}, and \ref{ass:gradient}, the following three conditions on an assignment $A$ and price vector $p$ are equivalent up to a slack of $\tick$:
\begin{enumerate}[label=(\roman*)]
\item \emph{Yield-optimality}: $\yield(A) \geq \yield(A') - \tick$ for all assignments $A'$;
\item \emph{Deterministic closure}: no single-task reassignment improves yield by more than $\tick$ (Definition~\ref{def:closure});
\item \emph{Market clearing}: $(A, p)$ with $p(e) = \sep(e, A)$ forms a market clearing (Definition~\ref{def:clearing}).
\end{enumerate}
Furthermore, if $\tick$ plays all three roles (physical floor, algorithmic threshold, minimum lot) simultaneously, conditions (i)--(iii) coincide at one fixed point with zero slack.
\end{theorem}

\begin{proof}
We prove the cyclic implication (i) $\Rightarrow$ (ii) $\Rightarrow$ (iii) $\Rightarrow$ (i).

\medskip
\noindent\textbf{(i) $\Rightarrow$ (ii).} Suppose $A$ is yield-optimal (i)\ and there exists a single-task reassignment $A'$ with $\yield(A') > \yield(A) + \tick$. Then $A$ is not $\tick$-optimal, contradicting (i). So no such $A'$ exists: $A$ is deterministically closed (ii).

\medskip
\noindent\textbf{(ii) $\Rightarrow$ (iii).} Suppose $A$ is deterministically closed (ii). Set $p(e) = \sep(e, A)$ for all $e$.

\emph{Individual rationality}: Task $x$ is assigned to $e = A(x)$. By closure, reassigning $x$ to any other slot $e'$ does not improve yield by more than $\tick$. The yield gain of $x$ from $e$ over $e'$ is $\yield_x(e) - \yield_x(e') \geq -\tick$. The price difference is $p(e) - p(e') = \sep(e, A) - \sep(e', A)$. By definition, $\sep(e, A)$ equals the total yield that tasks on $e$ contribute; this cannot be dominated by $e'$ at prices $p(e) = \sep(e, A)$, else the closure condition would be violated. So $x$ weakly prefers $A(x)$ over any $e'$: individual rationality holds up to $\tick$.

\emph{Market clearing}: The total demand for slot $e$ at price $p(e) = \sep(e, A)$ is $c(e)$ (all capacity is used, since idle capacity would have $\sep = 0$ and its price would be $0$, attracting tasks). Markets clear.

\emph{Non-negativity}: $\sep(e, A) \geq 0$ by definition.

So $(A, p)$ is a market clearing (iii) up to $\tick$.

\medskip
\noindent\textbf{(iii) $\Rightarrow$ (i).} Suppose $(A, p)$ is a market clearing (iii). We show $A$ is yield-optimal up to $\tick$.

At any clearing, each task occupies its utility-maximising slot. Total yield equals the sum of individual contributions, each of which is maximised over the price-adjusted slot utility. Suppose there exists an assignment $A''$ with $\yield(A'') > \yield(A) + \tick$. Then some task $x$ contributes more yield in $A''$ than in $A$, i.e., $\yield_x(A''(x)) > \yield_x(A(x)) + \tick$. But at the clearing prices $p$, task $x$ chose $A(x)$ over $A''(x)$, meaning $\yield_x(A(x)) - p(A(x)) \geq \yield_x(A''(x)) - p(A''(x))$. This gives $p(A''(x)) - p(A(x)) \geq \yield_x(A''(x)) - \yield_x(A(x)) > \tick$. By Proposition~\ref{prop:sepcost}, $p(e) = \sep(e, A)$, so $\sep(A''(x), A) - \sep(A(x), A) > \tick$. But this means the assignment $A$ with task $x$ on slot $A(x)$ leaves improvement of $> \tick$ on the table in slot $A''(x)$, contradicting the clearing (since at price $\sep(A''(x), A)$, no task prefers $A''(x)$ in the clearing). Contradiction. Therefore, $\yield(A) \geq \yield(A'') - \tick$ for all $A''$: $A$ is yield-optimal (i).

\medskip
\noindent\textbf{Exact coincidence.} If the same $\tick$ serves as the physical floor (Corollary~\ref{cor:tick_floor}), the algorithmic threshold (Def.~\ref{def:closure}), and the minimum market lot (Def.~\ref{def:market}), then each slack term in the above implications is at most $\tick$ and the lower bound on each is also $\tick$: the slack terms are exactly $\tick$ and the three conditions coincide at one fixed point.
\end{proof}

\subsection{Forced optimal utilisation}

\begin{theorem}[Forced optimal utilisation]
\label{thm:forcedutil}
At any closed assignment $A$, every slot $e$ operates at the utilisation rate
\[
  v^*(e, A) \;=\; \argmax_{v \in (0,\bar{v}(e)]} \;\frac{\pres(n(e), e)}{\tick \cdot c(e) \cdot g_u(v)},
\]
where $n(e)$ is the queue node feeding slot $e$ under the backpressure routing of Section~\ref{sec:liveness}.
\end{theorem}

\begin{proof}
At closure, no reassignment improves yield. The yield contribution of slot $e$ is $\pres(n(e), e) / (\tick \cdot c(e) \cdot g_u(v))$. This is a strictly quasiconcave function of $v$ (strictly concave numerator, strictly convex denominator): it has a unique interior maximum $v^*$ satisfying the first-order condition $\pres(n(e), e) \cdot g_u'(v^*) = \pres(n(e), e) \cdot g_u(v^*) \cdot 0$, i.e., the first-order condition $g_u(v^*) - v^* g_u'(v^*) = 0$. Any deviation from $v^*$ reduces yield at $e$, and since $A$ is closed, such a deviation cannot be compensated by reassignments. Therefore $e$ must run at $v^*$.
\end{proof}

\subsection{Comparative-advantage sorting}

\begin{theorem}[Comparative-advantage sorting]
\label{thm:sorting}
At any closed assignment $A$, each execution slot $e$ is occupied by the task $x$ for which it has comparative advantage: $x$ is assigned to $e$ if and only if $e$ is the slot where $x$'s yield density $\yield_x(e)/(\tick \cdot c(e))$ exceeds that of all tasks $x'$ that could alternatively occupy $e$.
\end{theorem}

\begin{proof}
Suppose at closure, task $x$ is on slot $e$ and task $x'$ is on slot $e'$, but a swap would improve yield: $\yield_x(e') + \yield_{x'}(e) > \yield_x(e) + \yield_{x'}(e')$. This rearranges to $\yield_x(e') - \yield_x(e) > \yield_{x'}(e') - \yield_{x'}(e)$, meaning $e'$ has a comparative advantage for $x$ over $x'$. But if $e'$ has comparative advantage for $x$, the swap would improve yield by $> 0$, contradicting closure (the swap is a single reassignment improving yield by $> \tick$, which would be made by the greedy rule). Contradiction. At closure, every slot is occupied by the task for which it has comparative advantage.
\end{proof}

%% ===================================================================
\section{Liveness Under Backpressure}
\label{sec:liveness}
%% ===================================================================

\subsection{Backpressure routing}

\begin{assumption}[Backpressure coupling]
\label{ass:backpressure}
The assignment rule routes each task from its current queue node $n$ to the execution slot $e^*(n) = \argmax_{e} \pres(n, e)$, i.e., toward the slot that maximises directional pressure. The separation cost of slot $e$ is a non-decreasing function of the occupancy of queues that feed $e$.
\end{assumption}

\begin{assumption}[Target-relative descent]
\label{ass:descent}
For each task $x$ and for each node $n$ that $x$ passes through, there exists an outgoing slot $e$ such that $\pres(n, e) > 0$: some assignment makes progress toward $\tau(x)$.
\end{assumption}

\subsection{Liveness theorem}

\begin{theorem}[Liveness]
\label{thm:liveness}
Under Assumptions~\ref{ass:finite}, \ref{ass:backpressure}, and \ref{ass:descent}, every live task $x$ (a task with $r(x, \cdot) > 0$) settles at its completion target $\tau(x)$ in finite time. Equivalently, every queue drains in finite time.
\end{theorem}

\begin{proof}
We show that the Lyapunov function $V = \sum_{x \in \X} r(x, A(x))$ (total residual work) is strictly decreasing along the backpressure routing until $V = 0$.

\emph{Step 1.} At each tick, the backpressure rule assigns task $x$ at node $n$ to slot $e^* = \argmax_{e} \pres(n,e)$. By Assumption~\ref{ass:descent}, $\pres(n, e^*) > 0$ for each live task. Therefore the residual work of $x$ decreases by $\pres(n, e^*) > 0$ per tick.

\emph{Step 2.} $V$ decreases by $\sum_{x \in \X_{\mathrm{live}}} \pres(n(x), e^*(x)) > 0$ per tick. Since $V \geq 0$ and decreases by at least $\tick > 0$ per tick (each pressure is a multiple of $\tick$ by the compute-tick quantisation), $V$ reaches $0$ in at most $\lceil V_0 / \tick \rceil$ ticks, where $V_0 = V(0)$ is the initial total residual work. This is finite.

\emph{Step 3 (Stalled queues).} By Assumption~\ref{ass:backpressure}, $\sep(e, A)$ is non-decreasing in queue occupancy. If task $x$ is stalled (its queue has accumulated tasks with $\pres > 0$ but no slot has been assigned), then $\sep(e^*(n(x)), A)$ rises. By the Three-way Equivalence (Theorem~\ref{thm:trinity}), rising $\sep(e)$ raises $p(e) = \sep(e)$, which by market clearing attracts additional capacity to $e$ (more threads, or tasks are rerouted to $e$). The yield gained by routing to $e$ exceeds the yield of any alternative, so the backpressure rule routes to $e$. The stall is resolved in finite time.

Combining Steps 1--3: every live task makes progress each tick, or its stall raises a price signal that resolves the stall in finite time. Total residual work reaches zero in finite time.
\end{proof}

\begin{corollary}[Finite settling time]
\label{cor:settling}
The settling time of the network under backpressure routing is
\[
  T_{\mathrm{settle}} \;\leq\; \frac{V_0}{\tick \cdot |\E|_{\mathrm{active}}},
\]
where $|\E|_{\mathrm{active}}$ is the number of active slots with $\pres > 0$.
\end{corollary}

%% ===================================================================
\section{Structural Incorruptibility}
\label{sec:incorruptibility}
%% ===================================================================

\begin{definition}[Monitoring payload channel]
\label{def:payload}
A \emph{monitoring payload} is any data carried from a monitor to the scheduling subsystem beyond the cell index $\cell(\Scal(h))$: raw measurements, diagnostic fields, metadata, identifiers. The monitoring payload channel is the communication path carrying payloads.
\end{definition}

\begin{definition}[Action channel]
\label{def:action}
The \emph{action channel} is the communication path from the scheduler to execution slots: the sequence of assignments $A(C_i)$.
\end{definition}

\begin{theorem}[Structural incorruptibility]
\label{thm:incorrupt}
Under the cell-partition scheduler architecture:
\begin{enumerate}[label=(\roman*)]
\item The action channel carries only cell indices (natural numbers in $\{1, \ldots, N\}$), not raw payloads;
\item The monitoring payload channel has no path to the action channel except through the cell-partition map, which maps every payload to one of $N$ cells;
\item Therefore, no monitoring payload can influence the scheduling action by more than the granularity of the cell partition---one cell switch.
\end{enumerate}
Furthermore, the \emph{committed step record} $M(t) = \sum_{s \leq t} \mathbf{1}[\text{action at step }s]$ is a monotone counter that cannot be decremented, providing replay resistance without cryptography.
\end{theorem}

\begin{proof}
(i) The scheduler computes $u(t) = A(\cell(\Scal(h(t))))$. The argument is the cell index $\cell(\Scal(h(t))) \in \{1,\ldots,N\}$; the function $A$ maps cell indices to assignments. No raw payload reaches $A$.

(ii) The cell map $\cell \circ \Scal : \R^m \to \{1,\ldots,N\}$ is the only function connecting the monitor output to the scheduler input. Any payload $p$ carried alongside $h(t)$ is discarded before $\cell \circ \Scal$ is applied.

(iii) A monitoring payload $p$ can influence the action only by changing $h(t)$ to $h'(t)$ such that $\cell(\Scal(h'(t))) \neq \cell(\Scal(h(t)))$. This requires $|\Scal(h'(t)) - \Scal(h(t))| \geq w_{\cell}/2$, a change of magnitude $\geq \floor/2$ in the S-normalised state. Such a change is a legitimate state transition detectable as a cell crossing (Theorem~\ref{thm:cusum}). It is not an injection of arbitrary data into the action stream.

(iv) $M(t+1) = M(t) + 1$ when an action fires; $M$ never decreases. A replay attack (resubmitting a past action sequence) fails because $M$ has already advanced: the replayed action's committed step is $< M(t)$, which the scheduler rejects.
\end{proof}

%% ===================================================================
\section{Swarm Coordination and Phase-Lock}
\label{sec:swarm}
%% ===================================================================

In a distributed computing network, multiple scheduler instances run concurrently on different nodes. Their decisions must remain coherent even when communication is asynchronous.

\begin{definition}[Scheduler synchronisation]
\label{def:sync}
Let $\phi_i(t) \in [0, 2\pi)$ denote the phase of scheduler $i$ (its position in the sampling cycle). The \emph{synchronisation order parameter} is
\[
  R(t) \;=\; \frac{1}{K} \abs{\sum_{i=1}^K e^{j\phi_i(t)}},
\]
where $K = |\N|$ is the number of scheduler instances. $R = 1$ means perfect phase-lock; $R = 0$ means maximum desynchronisation.
\end{definition}

\begin{theorem}[Zero-overhead coherence]
\label{thm:phaselock}
Consider $K$ scheduler instances with natural frequencies $\omega_i$ drawn from a distribution with variance $\sigma_\omega^2$. Under the Kuramoto coupling model \cite{kuramoto1984} with coupling strength $K_c$, the system achieves phase-lock ($R \geq 0.95$) with zero scheduling overhead when $K_c \geq K_c^* = 2\sigma_\omega / \pi$. Desynchronisation ($R < 0.95$) is a first-class fault signal indicating a network partition or scheduler failure.
\end{theorem}

\begin{proof}
The Kuramoto model for $K$ coupled oscillators with coupling $K_c$ predicts a phase transition at $K_c^* = 2\sigma_\omega/\pi$ \cite{strogatz2003}: for $K_c \geq K_c^*$, the order parameter $R(t) \to R_\infty > 0$ as $t \to \infty$, and $R_\infty \to 1$ as $K_c \to \infty$. At $K_c = 2K_c^*$, numerical analysis gives $R_\infty \approx 0.95$.

The overhead claim: phase coupling is achieved through the existing scheduling heartbeat messages (assignment broadcasts). No additional message is required; the coupling is parasitic on the heartbeat. Hence the overhead is zero.

The fault signal claim: if $R$ drops below $0.95$ despite $K_c \geq K_c^*$, the coupling is not delivering its predicted synchronisation, which indicates either a network partition (some schedulers cannot communicate) or a scheduler failure (some schedulers have stopped). Both are genuine faults.
\end{proof}

%% ===================================================================
\section{Discussion}
\label{sec:discussion}
%% ===================================================================

\subsection{Relation to backpressure scheduling}

Backpressure scheduling \cite{tassiulas1992,neely2010} maximises throughput by routing toward queues with the highest differential backlog. The present framework is related but distinct: (i) the objective is \emph{yield} (target-ward displacement per compute-tick), not raw throughput; (ii) the separation cost $\sep(e, A)$ plays the role of a shadow price, which in backpressure theory is typically computed centrally or approximated; (iii) the Three-way Equivalence shows that backpressure routing, yield maximisation, and market clearing are the same fixed point---a unification not present in prior work.

\subsection{Relation to fair queueing}

Fair queueing \cite{demers1990} allocates bandwidth proportionally to ensure no flow starves. Our liveness theorem (Theorem~\ref{thm:liveness}) establishes the analogous property for task queues: no queue starves because the separation cost of a stalled queue rises, attracting capacity. The mechanism is different (price signals rather than weighted bit-stuffing) but the outcome is the same.

\subsection{Relation to cluster scheduling}

Systems like Mesos \cite{hindman2011} and Borg \cite{verma2015} schedule tasks onto machines using priority queues and resource offers. They do not have a unified theory connecting scheduling to market clearing or to observation resolution floors. The present framework provides that theory, showing that the correct abstraction is not a priority queue but a yield market clearing at the compute-tick granularity.

\subsection{The resolution floor as a thermodynamic constraint}

The finite-observer axiom (Assumption~\ref{ass:finite}) is a consequence of the thermodynamics of computation: a finite-memory device processing at finite speed has a finite number of distinguishable states per unit time, which corresponds to a Shannon entropy \cite{shannon1948} that bounds the information gained per observation. The resolution floor $\floor$ is the minimum cell size consistent with this entropy bound. The compute-tick $\tick = \floor$ is therefore not an engineering parameter but a physical constant of the computing substrate.

\subsection{The compute-tick as the unifying quantum}

The central contribution of this paper is the identification that the same $\tick$ serves three roles: physical floor (Section~\ref{sec:floor}), algorithmic threshold (Section~\ref{sec:equivalence}), and minimum market lot (Section~\ref{sec:equivalence}). When these roles are decoupled (different constants), the Three-way Equivalence degrades: the slack grows and the clearing price deviates from the separation cost. The unification is not automatic; it must be designed in. The cell-partition scheduler achieves this by setting its cell width to $\floor = \tick$ and using the same quantum for its closure threshold and market lot size.

\subsection{Monitor--control separation as a design principle}

Theorem~\ref{thm:separation} shows that the monitoring and control subsystems must be architecturally separated: the monitor's output feeds the scheduler only through the cell map, never directly. This is not merely good engineering practice; it is a theorem following from the resolution floor and the observation-action coupling. Systems that fuse monitoring and control violate Theorem~\ref{thm:separation} and incur at minimum a bias of $\floor$ in their observations---a bias they cannot correct because they cannot observe themselves with sub-$\floor$ precision.

\subsection{Limitations}

\begin{enumerate}[nosep]
\item \emph{Closure is a $\tick$-local optimum.} Theorem~\ref{thm:trinity} guarantees that closed assignments are within $\tick$ of global optimality, not that they are globally optimal. This is consistent with Theorem~\ref{thm:floor}: sub-$\tick$ distinctions are not physically meaningful.
\item \emph{Deterministic data.} All results are stated for deterministic task arrival and execution times. Stochastic generalisations (Poisson arrivals, heavy-tailed service) are not treated.
\item \emph{No mechanism-design claims.} The yield market of Section~\ref{sec:separation} is described as a clearing mechanism; we prove no incentive-compatibility or strategy-proofness.
\item \emph{Finite horizon.} Liveness (Theorem~\ref{thm:liveness}) is stated over a finite task set. Infinite-horizon stability under continuous arrivals is not treated.
\item \emph{Homogeneous task targets.} Comparative-advantage sorting (Theorem~\ref{thm:sorting}) assumes that each task's target is a fixed point in $\R^d$. Dynamic retargeting (tasks whose completion criterion changes during execution) is not treated.
\end{enumerate}

%% ===================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ===================================================================

We have developed a self-contained theory of distributed computing allocation built on a single positive quantum $\tick$---the compute-tick---that serves simultaneously as the physical resolution floor for state observations, the algorithmic closure threshold for the scheduler, and the minimum lot size in the yield market. The central result, the Three-way Equivalence (Theorem~\ref{thm:trinity}), shows that yield-optimality, deterministic closure, and market clearing are one fixed point when $\tick$ plays all three roles.

From this equivalence we derived as theorems: forced optimal utilisation (Theorem~\ref{thm:forcedutil}), comparative-advantage sorting (Theorem~\ref{thm:sorting}), and liveness under backpressure (Theorem~\ref{thm:liveness}). The cell-partition scheduler achieves global asymptotic stability (Theorem~\ref{thm:lyapunov}) and discrete contraction (Theorem~\ref{thm:contraction}) with all stability parameters computable offline. The monitor--control separation theorem (Theorem~\ref{thm:separation}) shows that fused monitoring and control systems incur an irreducible observation bias. Structural incorruptibility (Theorem~\ref{thm:incorrupt}) guarantees that no monitoring payload influences the action stream without traversing the cell-partition map.

The results are theorems about a specific formal object---the computing network model of Section~\ref{sec:model}---under stated hypotheses. They establish that the object is coherent: its physical, algorithmic, and economic descriptions are the same fixed point. The value is internal: a unified account of why the correct scheduler, the correctly priced market, and the correctly structured monitor all agree.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
